### 2017-01-17 2017mergingmerges/section/01-Introduction.tex
mostly rejigging

Suppose we have two large files containing identifiers, and wish to perform some analysis on them.
Both files are sorted, but may contain duplicates.
We wish to retrieve the unique set from the first file, as well as the union of both files.
Can we perform both of these tasks at once, without reading the input files multiple times?

\begin{code}
uniquesUnion (file1 file2 : Stream Id) : (Stream Id, Stream Id)
 = let uniques = group file1
       merged  = merge file1 file
       union   = group merged
   in (uniques, union)
\end{code}

@@

@@

Fusion is then rather like executing two processes concurrently by taking the (synchronised) product of two state machines: either the both machines can take the same step (both pulling from @file1@, or one pushing to @merged@ while the other pulls), or one of the machines can take an independent step which isn't in the alphabet of the other while the other `rests'.
Deadlock can occur with synchronous product when each machine is waiting on the other, and neither can progress.
We solve this by also storing in the state label information about which input channels are known to have values.
This way, rather than requiring both machines to take the same step at the same time, executing a @push@ from one machine fills the buffer of the other, allowing the other machine to read the value from the buffer at its leisure.
In order to push, the other machine must have an empty buffer.
In this way we have extended synchronised product to allow more cases without deadlock.

We make the following contributions:
\begin{itemize}
\item a process calculus for encoding infinite streaming programs (\S\ref{s:Processes});
\item an algorithm for fusing processes, the first which supports splits, joins and arbitrary combinators (\S\ref{s:Fusion});
\item a formalisation and proof of soundness in Coq (\S\ref{s:Proofs});
\item an informal description of the required extensions to support finite streams (\S\ref{s:Finite});
\item and show that our calculus is general enough for many combinators, including segmented operations (\S\ref{s:Combinators}).
\end{itemize}

>>>>

Suppose we have two large files containing identifiers, and wish to perform some analysis on them.
Both files are sorted, but may contain duplicates.
We wish to retrieve the unique set from the first file, as well as the union of both files.
Can we perform both of these tasks at once, without reading the input files multiple times?
\TODO{More compelling justification. Why would we want to do this?}

\TODO{Should this example also have readFiles?}
\begin{code}
uniquesUnion file1 file2
 = do   let uniques = group file1
        let merged  = merge file1 file2
        let union   = group merged
        writeFile "uniques.txt" uniques
        writeFile "union.txt"   unions
\end{code}

% This example is somewhat incorrect because the streams here are pull streams,
% but the fused result cannot return pull streams (since we don't know what order they will be pulled)
% \begin{code}
% uniquesUnion (file1 file2 : Stream Id) : (Stream Id, Stream Id)
%  = let uniques = group file1
%        merged  = merge file1 file2
%        union   = group merged
%    in (uniques, union)
% \end{code}

@@

In order to fuse this example, we require both pull \emph{and} push streams.
The input streams must be pull streams since the order values are required is determined by the @merge@ combinator.
For the same reason, the outputs sent to each @writeFile@ must be push streams.

@@

\TODO{Should this be a concrete example of how they are fused?}
\TODO{Is the comparison to synchronised product helpful or just going to cause more confusion?}
Fusion is then rather like executing two processes concurrently by taking the (synchronised) product of two state machines: either the both machines can take the same step (both pulling from @file1@, or one pushing to @merged@ while the other pulls), or one of the machines can take an independent step which isn't in the alphabet of the other while the other `rests'.
Deadlock can occur with synchronous product when each machine is waiting on the other, and neither can progress.
We solve this by also storing in the state label information about which input channels are known to have values.
This way, rather than requiring both machines to take the same step at the same time, executing a @push@ from one machine fills the buffer of the other, allowing the other machine to read the value from the buffer at its leisure.
In order to push, the other machine must have an empty buffer.
In this way we have extended synchronised product to allow more cases without deadlock.

We make the following contributions:
\begin{itemize}
\item a process calculus for encoding infinite streaming programs (\S\ref{s:Processes});
\item an algorithm for fusing processes, the first which supports splits, joins and arbitrary combinators (\S\ref{s:Fusion});
\item a formalisation and proof of soundness in Coq (\S\ref{s:Proofs});
\item an informal description of the required extensions to support finite streams (\S\ref{s:Finite});
\item and show that our calculus is general enough for many combinators, including segmented operations (\S\ref{s:Combinators}).
  \TODO{Is this list of combinators helpful or just boring?}
\end{itemize}

### 2017-01-17 2017mergingmerges/section/A1-Combinators.tex
mostly rejigging

Here we show the definitions of some combinators.
We start with simple combinators supported by most streaming systems, and progress to more interesting combinators.
Some combinators such as @fold@, @take@ and @append@ are missing due to the infinite nature of our streams, but can be implemented in our finite stream extension described in~\S\REF.
The fact that segmented versions of these combinators can be implemented is compelling evidence of this.

@@

@@

\begin{code}
merge = process
     ins: as bs
    outs: abs
    heap: {a, b}
   label: pI0
  blocks: pI0 = pull as a  pI1
          pI1 = pull bs b  pC0
          pC0 = case (a < b) pA0 pB0

          pA0 = push abs a pA1
          pA1 = drop as    pA2
          pA2 = pull as  a pC0

          pB0 = push abs b pB1
          pB1 = drop bs    pB2
          pB2 = pull bs  b pC0
\end{code}

>>>>

Here we show the definitions of some combinators.
We start with simple combinators supported by most streaming systems, and progress to more interesting combinators.
Some combinators such as @fold@, @take@ and @append@ are missing due to the infinite nature of our streams, but can be implemented in our finite stream extension described in~\S\ref{s:Finite}.
The fact that segmented versions of these combinators can be implemented is compelling evidence of this.

@@

\paragraph{Group} removes filters out consecutive duplicates from a stream.
If the stream is sorted, the output will contain only the unique entries in the stream.

\begin{code}
group = process
     ins: as
    outs: bs
    heap: {last, val}
   label: I0
  blocks: I0 = pull as val I1
          I1 = push bs val I2[last = val]
          I2 = drop as     R0

          R0 = pull as val R1
          R1 = case (last /= val) R2 R3
          R2 = push bs val R3[last = val]
          R3 = drop as     R0
\end{code}

@@

\begin{code}
merge = process
     ins: as bs
    outs: abs
    heap: {a, b}
   label: I0
  blocks: I0 = pull as a  I1
          I1 = pull bs b  C0
          C0 = case (a < b) A0 B0

          A0 = push abs a A1
          A1 = drop as    A2
          A2 = pull as  a C0

          B0 = push abs b B1
          B1 = drop bs    B2
          B2 = pull bs  b C0
\end{code}

### 2017-01-17 2017mergingmerges/section/XX-Processes.tex
mostly rejigging

A process defines a stream computation, taking any number of input streams and producing at least one output stream.
The process name is just for fun.
The input streams are paired with an input state which is used for coordinating multiple processes during evaluation; in process definitions before any evaluation has occurred this should be @none@.
The input states are explained in detail later in \S\ref{s:Process:Eval}.

@@

\begin{code}
map f = process
    name: (map f)
     ins: as
    outs: bs
    heap: {a = 0}
   label: map0
  blocks: map0 = pull as    a  map1
          map1 = push bs (f a) map2
          map2 = drop as       map0
\end{code}

@@

\begin{code}
process
    name: (map f)
     ins: xs
    outs: ys
    heap: {x}
   label: p0
  blocks: p0 = pull xs    x  p1
          p1 = push ys (f x) p2
          p2 = drop xs       p0
process
    name: (map g)
     ins: ys
    outs: zs
    heap: {y}
   label: q0
  blocks: q0 = pull ys    y  q1
          q1 = push zs (g y) q2
          q2 = drop ys       q0
\end{code}

@@

\begin{code}
process
    name: (map f / map g)
     ins: xs
    outs: ys zs
    heap: {x, y, _ys}
   label: p0q0
  blocks: p0q0            = pull xs    x  p1q0
          p1q0            = push ys (f x) p2q0-pending-ys { _ys = f x }
          p2q0-pending-ys = drop xs       p0q0-pending-ys
          p0q0-pending-ys = jump          p0q1-have-ys    { y = _ys }
          p0q1-have-ys    = push zs (g y) p0q2-have-ys
          p0q2-have-ys    = jump          p0q0
\end{code}

>>>>

A process defines a stream computation, taking any number of input streams and producing at least one output stream.
Processes have an optional name in parentheses used only for decsriptive purposes, and does not need to be unique.
The input streams are paired with an input state which is used for coordinating multiple processes during evaluation; in process definitions before any evaluation has occurred this should be @none@.
The input states are explained in detail later in \S\ref{s:Process:Eval}.

@@

\begin{code}
map f = process (map f)
     ins: as
    outs: bs
    heap: {a = 0}
   label: map0
  blocks: map0 = pull as    a  map1
          map1 = push bs (f a) map2
          map2 = drop as       map0
\end{code}

@@

\begin{code}
process (map f)
     ins: xs
    outs: ys
    heap: {x}
   label: p0
  blocks: p0 = pull xs    x  p1
          p1 = push ys (f x) p2
          p2 = drop xs       p0
process (map g)
     ins: ys
    outs: zs
    heap: {y}
   label: q0
  blocks: q0 = pull ys    y  q1
          q1 = push zs (g y) q2
          q2 = drop ys       q0
\end{code}

@@

\begin{code}
process (map f / map g)
     ins: xs
    outs: ys zs
    heap: {x, y, _ys}
   label: p0q0
  blocks: p0q0            = pull xs    x  p1q0
          p1q0            = push ys (f x) p2q0-pending-ys { _ys = f x }
          p2q0-pending-ys = drop xs       p0q0-pending-ys
          p0q0-pending-ys = jump          p0q1-have-ys    { y = _ys }
          p0q1-have-ys    = push zs (g y) p0q2-have-ys
          p0q2-have-ys    = jump          p0q0
\end{code}
